---
title: "R Notebook on 9.3_clonotype_top100_bubbleplot"
output: html_notebook
---
#### Get the input file:
After the "4_Normalization_clustering_BasicViz" step, there should be a file named "data_clust_cc_stress_regressed.rds". Copy this file to the current folder where you have this .rmd file ("5_cluster_freq_calculation").

#### Load required packages:
To load the required packages, execute the following code:

```{r load Libraries}
library(Seurat)
library(ggplot2)
library(readxl)
library(tibble)
library(dplyr)
library(patchwork)
library(ggrepel)
library(RColorBrewer)
library(tidyr)
library(openxlsx)
library(packcircles)
library(viridis)
```

Note: If you haven't installed the packages yet, please install them before loading.

#### Step 1: Data Loading and adding len_g and len_d columsn
Read the clustered data by running the following code:

```{r}
data_clust <- readRDS("data_clust_cc_regressed.rds")

# Let's take a look at the data to ensure everything looks good
DimPlot(data_clust, reduction = "umap", label = TRUE) + NoLegend()
```

#### Step 2: Make bubble plots for specific or all TRGV clonotypes while color coding TRDV usage
In this section, TRGV specific variables are declared to subset specific cell types or specific clusters or specific TRGV+ cells. Also, the variables are used to generate the final file name automatically. 
Moreover, here the color coding is for separating specific TRDV+ clonotypes from each other.
```{r}
organ.id <- "katherine"  # 'organ.id' represents the identifier for the specific organ or sample used in the analysis, set as "E17".

# Arrays specifying all available cell types and clusters for subsequent analysis and naming conventions
hash.all <- levels(data_clust@meta.data$hash.type)
cluster.all <- levels(data_clust@meta.data$seurat_clusters)
trgv.all <- levels(data_clust@meta.data$v_gene_g)


# Define the specific hash.type clusters to analyze, either "B6" or "SAP". Adjust 'HASHTYPE.TITLE' accordingly.
HASHTYPE <- "KS" #Default = hash.all
HASHTYPE.TITLE <- "KS"


# Define the specific clusters to analyze, either "0" or a specific set or use all available clusters 'cluster.all'. Adjust 'SEURAT_CLUSTERS.TITLE' accordingly.
SEURAT_CLUSTERS <- cluster.all #Default = cluster.all
SEURAT_CLUSTERS.TITLE <- "Clusterall"

#Define which TCRγ chain (v_gene_g) you wanna look at - "TRGV2" or "TRGV4" or trgv.all
V_GENE_G <- "TRGV4"
V_GENE_G.TITLE <- "trgv4"

# Define the number of dots (unique clonotypes) to be shown. Here, the default is 100. 
DOTS <- 50
DOTS_Title <- "50"
```

Now, run a common script based on the variables set above
```{r}
# Removal of "NA" from both gamma and delta chain important for this script to work
# Replace NA values in v_gene_g with "NA"
data_clust@meta.data$v_gene_g[is.na(data_clust@meta.data$v_gene_g)] <- "NA"
data_clust@meta.data$v_gene_d[is.na(data_clust@meta.data$v_gene_d)] <- "NA"
data_clust@meta.data$cdr3_g[is.na(data_clust@meta.data$cdr3_g)] <- "NA"
data_clust@meta.data$cdr3_d[is.na(data_clust@meta.data$cdr3_d)] <- "NA"
data_clust@meta.data$j_gene_g[is.na(data_clust@meta.data$j_gene_g)] <- "NA"
data_clust@meta.data$d_gene_d[is.na(data_clust@meta.data$d_gene_d)] <- "NA"
data_clust@meta.data$j_gene_d[is.na(data_clust@meta.data$j_gene_d)] <- "NA"

sum(is.na(data_clust@meta.data$v_gene_g))
sum(is.na(data_clust@meta.data$v_gene_d))
sum(is.na(data_clust@meta.data$cdr3_g))
sum(is.na(data_clust@meta.data$cdr3_d))

#Set up levels of TRDV genes
data_clust$v_gene_d <- factor(data_clust$v_gene_d, levels =c(c("TRAV12-3", "TRAV13-4/DV7", "TRAV13-5", "TRAV13D-1", "TRAV13D-2", "TRAV13D-3", 
  "TRAV13D-4", "TRAV13N-2", "TRAV14-1", "TRAV14-2", "TRAV14D-3/DV8", 
  "TRAV15-1/DV6-1", "TRAV15D-2/DV6D-2", "TRAV15N-1", "TRAV16N", "TRAV21/DV12", 
  "TRAV23", "TRAV3-4", "TRAV5D-4", "TRAV6D-7", "TRAV7-6", "TRDV1", "TRDV2-2", 
  "TRDV4", "TRDV5", "NA")))

Idents(data_clust) <- "v_gene_d"
data.subset <- subset(data_clust, idents = "NA", invert=TRUE)

Idents(data.subset) <- "v_gene_g"
data.subset <- subset(data.subset, idents = V_GENE_G) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "cdr3_g"
#data.subset <- subset(data.subset, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "cdr3_d"
data.subset <- subset(data.subset, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "hash.type"
data.subset <- subset(data.subset, idents = HASHTYPE)

#Idents(data.subset) <- "seurat_clusters"
#data.subset <- subset(data.subset, idents = SEURAT_CLUSTERS) # (If there is any cluster missing in the subsetted data, this line will not work. in that case just skip the line)
data.subset.vdj <- as.data.frame(data.subset@meta.data)

#Now group based on Unique CDR3s
data_group <- data.subset.vdj%>%                   
  group_by(.dots=c("v_gene_g","j_gene_g","cdr3_g","cdr3_d","v_gene_d","d_gene_d","j_gene_d","seurat_clusters")) %>%        
  summarise(count = n())%>%
  arrange(desc(count))
#View(data_group)

  
#Now convert "new_clusts" id like this - 7 becomes C7. Then add count of each clonotype in parenthesis - C7(5)
data_group_modified <- data_group%>% 
  mutate(C = paste("C"))%>%
  unite(new_clusts_ID, c("C", "seurat_clusters"), sep = "")%>%
  mutate(new_clusts_count=paste(new_clusts_ID, "(", count,")", sep =""))
#View(data_group_modified) 


#Now merge unique clonotypes from different clusters
data_group_final <- data_group_modified%>%                   
  group_by(.dots=c("v_gene_g","j_gene_g","cdr3_g","cdr3_d","v_gene_d","d_gene_d","j_gene_d")) %>%        
  summarise(Count = sum(count), cluster_ID_w_count=paste(new_clusts_count, collapse=","))

  #Shuffle the rows so that singletons are not ordered by v_gene_g usage
data_group_final2 <- data_group_final[sample(1:nrow(data_group_final)), ]

data_group_final3 <- data_group_final2%>%
  arrange(desc(Count))%>%
  mutate(Clonotype=paste0(v_gene_g,"-",cdr3_g,"-",cdr3_d,"-", v_gene_d))

# Sum count of all the top 100 clonotypes
sum_cell <- sum(data_group_final3$Count)

# Add frequency column
data_group_freq <- data_group_final3%>%                   
  mutate(freq = Count/sum_cell)

# Keep the top 100 or 50 or 47(DOTS) clones. Update the DOTs value above to make sure it does not exceed the total number of clonotypes found in the subset
data_group_DOTS <- head(data_group_freq, DOTS)
data_group_DOTS_save <- tibble::rownames_to_column(data_group_DOTS, "Clonotype_ID")
write.xlsx(as.data.frame(data_group_DOTS_save),paste0(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_G.TITLE,"-",SEURAT_CLUSTERS.TITLE,".xlsx", sep = ""), row.names=FALSE, col.names = TRUE, append = FALSE)


# Generate the layout. sizetype can be area or radius, following your preference on what to be proportional to value.
packing <- circleProgressiveLayout(data_group_DOTS$freq, sizetype='area')
data_group_pack <- cbind(data_group_DOTS, packing)
dat.gg <- circleLayoutVertices(packing, npoints=DOTS)
packing2 <- cbind(packing, data_group_pack$v_gene_d)
# change column name
packing2 <- data.frame(
  x = packing$x,
  y = packing$y,
  radius = packing$radius,
  v_gene_d = data_group_pack$v_gene_d
)
#Make the v_gene_d Values Unique (this is to solve the issue with the "circleLayoutVertices" argument in the following step)
packing2 <- packing2 %>%
  group_by(v_gene_d) %>%
  mutate(v_gene_d = paste0(v_gene_d, "_", row_number())) %>%
  ungroup()

dat.gg2 <- circleLayoutVertices(packing2, npoints=DOTS, idcol = 4)
# add v_gene_g info the the dat.gg file. This will help with coloring
dat.gg.bind <- cbind(dat.gg, dat.gg2$id)
#rename column names
colnames(dat.gg.bind) <- c('x','y','id','v_gene_d')

# remove the unwanted part ("_") of each string in the v_gene_d column that I introduced to make the values unique
dat.gg.bind$v_gene_d <- sub("_\\d+$", "", dat.gg.bind$v_gene_d)

# Set up color code for TRDV chains
chainname <- c(chainname <- c(
  "TRDV1", "TRDV2-2", "TRDV4", "TRDV5", 
  "TRAV12-3", "TRAV13-4/DV7", "TRAV13-5", "TRAV13D-1", "TRAV13D-2", "TRAV13D-3", 
  "TRAV13D-4", "TRAV13N-2", "TRAV14-1", "TRAV14-2", "TRAV14D-3/DV8", 
  "TRAV15-1/DV6-1", "TRAV15D-2/DV6D-2", "TRAV15N-1", "TRAV16N", "TRAV21/DV12", 
  "TRAV23", "TRAV3-4", "TRAV5D-4", "TRAV6D-7", "TRAV7-6", "NA"
)) 
colorHex <- c(colorHex <- c(
  "#fddbc7", "#c2a5cf", "#de77ae", "#fdae61", "#c7eae5", "#ffffbf", 
  "#e6f5d0", "#bababa", "#8c510a", "#003c30", "#80cdc1", "#7fbc41", 
  "#f7f7f7", 
  "#b3e2cd", "#fdcdac", "#cbd5e8", "#f4cae4", "#e6f598", "#abd9e9", 
  "#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854", "#ffd92f","#1a1a1a"
)) # This is the the chosen color scheme for each of the expected TRDV chains. Edit this color scheme and order to fit the total number of unique TRDV chains identified

colorbank <- data.frame(chainname,colorHex)

# Identify the TRDV chains that are only present in the subsetted datset
subset_TRDV <- data_group_DOTS%>%                   
  group_by(.dots=c("v_gene_d"))%>%
  summarise(count = n())
# Merge two dataframes together containing common characters in specific columns
color_select <- left_join(subset_TRDV, colorbank, by=c('v_gene_d'='chainname'))

# Basic color customization
ggplot() + 
  geom_polygon(data = dat.gg.bind, aes(x, y, group = id, fill=as.factor(v_gene_d)), colour = "black", alpha = 0.9) +
  scale_fill_manual(values = color_select$colorHex) +
  labs(title = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_G.TITLE,"-",SEURAT_CLUSTERS.TITLE, sep = ""))+
  scale_size_continuous(range = c(1,4)) +
  theme_void() + 
  theme(legend.position="right") +
  coord_equal()

ggsave(path = "Figs", filename = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_G.TITLE,"-",SEURAT_CLUSTERS.TITLE,"-",DOTS_Title,"clones.png", sep = ""),  height=5, width=8, units='in', dpi = 300, bg = "transparent", device='png')

ggsave(path = "Figs", filename = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_G.TITLE,"-",SEURAT_CLUSTERS.TITLE,"-",DOTS_Title,"clones.pdf", sep = ""),  height=5, width=8, units='in', dpi = 300, bg = "transparent", device='pdf')

#################
# Note: The actual sequences of the top clonotypes of the bubble plot can be inferred from the .xlsx file of the same name that this script automatically produces
#################
```


#### Step 3: Make bubble plots for specific or all TRDV clonotypes while color coding TRGV usage
In this section, TRGV specific variables are declared to subset specific cell types or specific clusters or specific TRGV+ cells. Also, the variables are used to generate the final file name automatically. 
Moreover, here the color coding is for separating specific TRDV+ clonotypes from each other.
```{r}
organ.id <- "E17"  # 'organ.id' represents the identifier for the specific organ or sample used in the analysis, set as "E17".

# Arrays specifying all available cell types and clusters for subsequent analysis and naming conventions
hash.all <- levels(data_clust@meta.data$hash.type)
cluster.all <- levels(data_clust@meta.data$seurat_clusters)
trdv.all <- levels(data_clust@meta.data$v_gene_d)


# Define the specific hash.type clusters to analyze, either "B6" or "SAP". Adjust 'HASHTYPE.TITLE' accordingly.
HASHTYPE <- "B6" #Default = hash.all
HASHTYPE.TITLE <- "B6"


# Define the specific clusters to analyze, either "0" or a specific set or use all available clusters 'cluster.all'. Adjust 'SEURAT_CLUSTERS.TITLE' accordingly.
SEURAT_CLUSTERS <- cluster.all #Default = cluster.all
SEURAT_CLUSTERS.TITLE <- "Clusterall"

#Define which TCRγ chain (v_gene_g) you wanna look at - "TRGV2" or "TRGV4" or trgv.all
V_GENE_D <- trdv.all
V_GENE_D.TITLE <- "trdv.all"

# Define the number of dots (unique clonotypes) to be shown. Here, the default is 100. 
DOTS <- 100
DOTS_Title <- "100"
```

Now, run a common script based on the variables set above
```{r}
#Set up levels of TRGV genes
data_clust$v_gene_g <- factor(data_clust$v_gene_g, levels =c("TRGV1","TRGV2","TRGV3","TRGV4","TRGV5","TRGV6","TRGV7","NA"))

#Removal of "NA" from both gamma and delta chain important for this script to work
Idents(data_clust) <- "org_v_gene_d" 
data.subset <- subset(data_clust, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "v_gene_g"
data.subset <- subset(data.subset, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "cdr3_g"
data.subset <- subset(data.subset, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)

Idents(data.subset) <- "cdr3_d"
data.subset <- subset(data.subset, idents = "NA", invert=TRUE) # (If there is no "NA", please skip this line)


#Make specific subset based on the variables set in the previous step
Idents(data.subset) <- "hash.type"
data.subset <- subset(data.subset, idents = HASHTYPE)

Idents(data.subset) <- "seurat_clusters"
data.subset <- subset(data.subset, idents = SEURAT_CLUSTERS) # (If there is any cluster missing in the subsetted data, this line will not work. in that case just skip the line)
data.subset.vdj <- as.data.frame(data.subset@meta.data)

#Now group based on Unique CDR3s
data_group <- data.subset.vdj%>%                   
  group_by(.dots=c("v_gene_g","j_gene_g","cdr3_g","cdr3_d","org_v_gene_d","d_gene_d","j_gene_d","seurat_clusters")) %>%        
  summarise(count = n())%>%
  arrange(desc(count))
#View(data_group)

  
#Now convert "new_clusts" id like this - 7 becomes C7. Then add count of each clonotype in parenthesis - C7(5)
data_group_modified <- data_group%>% 
  mutate(C = paste("C"))%>%
  unite(new_clusts_ID, c("C", "seurat_clusters"), sep = "")%>%
  mutate(new_clusts_count=paste(new_clusts_ID, "(", count,")", sep =""))
#View(data_group_modified) 


#Now merge unique clonotypes from different clusters
data_group_final <- data_group_modified%>%                   
  group_by(.dots=c("v_gene_g","j_gene_g","cdr3_g","cdr3_d","org_v_gene_d","d_gene_d","j_gene_d")) %>%        
  summarise(Count = sum(count), cluster_ID_w_count=paste(new_clusts_count, collapse=","))

  #Shuffle the rows so that singletons are not ordered by v_gene_g usage
data_group_final2 <- data_group_final[sample(1:nrow(data_group_final)), ]

data_group_final3 <- data_group_final2%>%
  arrange(desc(Count))%>%
  mutate(Clonotype=paste0(v_gene_g,"-",cdr3_g,"-",cdr3_d,"-", org_v_gene_d))

# Sum count of all the top 100 clonotypes
sum_cell <- sum(data_group_final3$Count)

# Add frequency column
data_group_freq <- data_group_final3%>%                   
  mutate(freq = Count/sum_cell)


# Keep the top 100 clones 
data_group_100 <- head(data_group_freq, 100)
data_group_100_save <- tibble::rownames_to_column(data_group_100, "Clonotype_ID")
write.xlsx(as.data.frame(data_group_DOTS_save),paste0(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_D.TITLE,"-",SEURAT_CLUSTERS.TITLE,".xlsx", sep = ""), row.names=FALSE, col.names = TRUE, append = FALSE)


# Generate the layout. sizetype can be area or radius, following your preference on what to be proportional to value.
packing <- circleProgressiveLayout(data_group_100$freq, sizetype='area')
data_group_pack <- cbind(data_group_100, packing)
dat.gg <- circleLayoutVertices(packing, npoints=100)
packing2 <- cbind(packing, data_group_pack$v_gene_g)
dat.gg2 <- circleLayoutVertices(packing2, npoints=100, idcol = 4)
# add v_gene_g info the the dat.gg file. This will help with coloring
dat.gg.bind <- cbind(dat.gg, dat.gg2$id)
#rename column names
colnames(dat.gg.bind) <- c('x','y','id','v_gene_g')


# Basic color customization
ggplot() + 
  geom_polygon(data = dat.gg.bind, aes(x, y, group = id, fill=as.factor(v_gene_g)), colour = "black", alpha = 0.9) +
  # Adjust colors based on number of v_gene_g chains. default: c("#66c2a5","#ffffb3","#bebada","#fb8072","#80b1d3","#fdb462","#b3de69")
  scale_fill_manual(values = c("#66c2a5","#ffffb3","#bebada","#fb8072","#80b1d3","#fdb462","#b3de69")) +
  labs(title = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_D.TITLE,"-",SEURAT_CLUSTERS.TITLE, sep = ""))+
  scale_size_continuous(range = c(1,4)) +
  theme_void() + 
  theme(legend.position="right") +
  coord_equal()

ggsave(path = "Figs", filename = paste(organ.id,"-",HASHTYPE.TITLE,"-",SEURAT_CLUSTERS.TITLE,".png", sep = ""),  height=5, width=8, units='in', dpi = 300, bg = "transparent", device='png')

ggsave(path = "Figs", filename = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_D.TITLE,"-",SEURAT_CLUSTERS.TITLE,"-",DOTS_Title,"clones.png", sep = ""),  height=5, width=8, units='in', dpi = 300, bg = "transparent", device='png')
#ggsave(path = "Figs", filename = paste(organ.id,"-",HASHTYPE.TITLE,"-",V_GENE_D.TITLE,"-",SEURAT_CLUSTERS.TITLE,"-",DOTS_Title,"clones.svg", sep = ""),  height=5, width=8, units='in', dpi = 300, bg = "transparent", device='svg')


#################
# Note: The actual sequences of the top clonotypes of the bubble plot can be inferred from the .xlsx file of the same name that this script automatically produces
#################
```
