---
title: "R Notebook on 9.4_clonotype_featureplot"
output: html_notebook
---
#### Get the input file:
After the "4_Normalization_clustering_BasicViz" step, there should be a file named "data_clust_cc_stress_regressed.rds". Copy this file to the current folder where you have this .rmd file ("5_cluster_freq_calculation").

#### Load required packages:
To load the required packages, execute the following code:
```{r}
library(Seurat)
library(ggplot2)
library(readxl)
library(tibble)
library(dplyr)
library(patchwork)
library(ggrepel)
library(RColorBrewer)
library(tidyr)
library(ggpubr)
```

Note: If you haven't installed the packages yet, please install them before loading.

#### Step 1: Data Loading and creating new column named "clonotype"
Read the clustered data by running the following code:
```{r}
data_clust <- readRDS("data_clust_cc_stress_regressed.rds")

# Let's take a look at the data to ensure everything looks good
DimPlot(data_clust, reduction = "umap", label = TRUE) + NoLegend()
```

Now make a new column named clonotype in the following format :
TRGV-cdr3_g-cdr3_d-TRDV: TRGV5-CACWDSSGFHKVF-CASGYIGGIRATDKLVF-TRAV13-4/DV7

```{r}
data_clust@meta.data$clonotype <- paste(data_clust@meta.data$v_gene_g, data_clust@meta.data$cdr3_g, data_clust@meta.data$cdr3_d, data_clust@meta.data$org_v_gene_d, sep="-")
```

#### Step 2: Set variables and run common script
```{r}
organ.id <- "E17_thy"  # 'organ.id' represents the identifier for the specific organ or sample used in the analysis, set as "E17".

# Arrays specifying all available cell types and clusters for subsequent analysis and naming conventions
hash.all <- levels(data_clust@meta.data$hash.type)
cluster.all <- levels(data_clust@meta.data$seurat_clusters)
trgv.all <- levels(data_clust@meta.data$v_gene_g)


# Define the specific hash.type clusters to analyze, either "B6" or "SAP". Adjust 'HASHTYPE.TITLE' accordingly.
HASHTYPE <- "B6" #Default = hash.all
HASHTYPE.TITLE <- "B6"


# Define the specific clusters to analyze, either "0" or a specific set or use all available clusters 'cluster.all'. Adjust 'SEURAT_CLUSTERS.TITLE' accordingly.
SEURAT_CLUSTERS <- cluster.all #Default = cluster.all
SEURAT_CLUSTERS.TITLE <- "Clusterall"

# Define clonotypes to visualize
CLONOTYPE <- c("TRGV5-CACWDSSGFHKVF-CGSDIGGSSWDTRQMFF-TRDV4","TRGV6-CACWDSSGFHKVF-CGSDIGGSSWDTRQMFF-TRDV4","TRGV1-CAVWIGTSWVKIF-CGSDIGGSSWDTRQMFF-TRDV4","TRGV4-CSYGYSSGFHKVF-CASGYIGGIRATDKLVF-TRDV5","TRGV4-CSYGYSSGFHKVF-CASGYRRDTDKLVF-TRDV5","TRGV4-CSYGYSSGFHKVF-CALMERGIRATDKLVF-TRDV2-2")

# Set number of columns and Rows for final figure
NCOL <- 3
NROW <- 2

# Set final figure size in inches (based on how many clonotypes we want to look at once)
HEIGHT <- 13
WIDTH <- 24
```

Now, run a common script based on the variables set above
```{r}
# Make specific subset based on the variables set in the previous step
# Subset the data based on the selected hash type
Idents(data_clust) <- "hash.type"
data.subset <- subset(data_clust, idents = HASHTYPE)

# Subset the data further based on the selected Seurat clusters
Idents(data.subset) <- "seurat_clusters"
data.subset <- subset(data.subset, idents = SEURAT_CLUSTERS)

# Create an empty list to store the plots.
plot_list <- list()

# Loop through the clonotypes.
for (i in seq_along(CLONOTYPE)) {
  # Create a new column in metadata based on the matching condition
  data.subset@meta.data <- data.subset@meta.data %>%
    mutate(clonotype_renamed = ifelse(clonotype == CLONOTYPE[i], clonotype, "Others"))
  # Identify the unique clonotypes in the subset
  Idents(data.subset) <- "clonotype_renamed"
  # Generate a dimensionality reduction plot for the current clonotype
  plot <- DimPlot(data.subset, reduction = "umap", cols = c("grey", "blue"), label = FALSE, pt.size = 3, order = c(CLONOTYPE[i], "Others")) +
    ggtitle(CLONOTYPE[i]) + NoLegend()
  # Store the plot in the list
  plot_list[[i]] <- plot
}

# Display the generated plots.
for (i in seq_along(CLONOTYPE)) {
  print(plot_list[[i]])
}

# Arrange the plots into a single figure
figure <- ggarrange(plotlist=plot_list,
                    hjust = -1.8, vjust = 1.1,
                    ncol = NCOL, nrow = NROW)

# Save the figure as a PNG file
ggsave(figure, path="Figs", filename = paste(organ.id,"-", HASHTYPE.TITLE,"-",SEURAT_CLUSTERS.TITLE,".png", sep = ""),  height=HEIGHT, width=WIDTH, units='in', dpi = 300, bg = "transparent", device='png')

# Save the figure as an SVG file
#ggsave(figure, path="Figs", filename = paste(organ.id,"-", HASHTYPE.TITLE,"-",SEURAT_CLUSTERS.TITLE,".svg", sep = ""),  height=HEIGHT, width=WIDTH, units='in', dpi = 300, bg = "transparent", device='svg')
```

