---
title: "R Notebook - Custom VDJ Processing - Cellranger to Seurat compatible format"
output: html_notebook
---

#### Load required packages:
To load the required packages, execute the following code:

```{r}
library(tidyverse)
library(Matrix)
library(RCurl)
library(openxlsx)
library(here)
library(dplyr)
library(writexl)
```

Note: If you haven't installed the packages yet, please install them before loading.

#### Data loading
Load barcode.tsv.gz file from "cellranger count" output and "filtered_contig_annotations.csv" file (for ab T cells) or "raw_contig_annotations.csv" file (for gd T cells)
```{r}
# Read the .tsv.gz file that contains barcodes from the cellranger count output
barcodes <- read.delim(gzfile(here("data", "barcodes.tsv.gz")), header = FALSE, stringsAsFactors = FALSE)
colnames(barcodes) <- c("barcode")

# Read the CSV file
data <- read.csv(here("data", "filtered_contig_annotations.csv"), header = TRUE, stringsAsFactors = FALSE)
```

#### Data procesing and splitting into TRA or TRB 
(Optional) if processing ab T cells, Filter out rows where 'productive' is "true". If processing gd T cells, DO NOT run this line!
```{r}
# Filter out rows where 'productive' is TRUE
data <- data %>%
  filter(productive == "true")
```


```{r}
# Split the filtered data into TRA and TRB
data_TRA <- data %>%
  filter(chain == "TRA")

data_TRB <- data %>%
  filter(chain == "TRB")

# Rename columns in data_TRA while keeping 'barcode' unchanged
data_TRA <- data_TRA %>%
  rename_with(~ paste0(., "_a"), -barcode)  # Exclude 'barcode' from renaming

# Rename columns in data_TRB while keeping 'barcode' unchanged
data_TRB <- data_TRB %>%
  rename_with(~ paste0(., "_b"), -barcode)  # Exclude 'barcode' from renaming
```

#### processing of data_TRA

```{r}
# Initialize an empty data frame to store the output
data_TRA_output <- data.frame()

# Get unique barcodes in the order they appear
unique_barcodes <- unique(data_TRA$barcode)

# Loop through each unique barcode
for (barcode in unique_barcodes) {
  # Filter the data for the current barcode
  current_data <- data_TRA %>%
    filter(barcode == !!barcode)

  # Count the number of rows for the current barcode, excluding the header
  total_contigs_a <- nrow(current_data)

  # Sum the reads column for the current barcode
  sum_reads <- sum(current_data$reads_a)

  # Calculate frequency of the current barcode
  contig_freq_a <- current_data$reads_a / sum_reads

  # Create a new data frame for the current barcode and move total_contigs column
  current_output <- current_data %>%
    mutate(total_contigs_a = total_contigs_a,
           contig_freq_a = contig_freq_a) %>%
    select(barcode, total_contigs_a, contig_freq_a, everything()) %>%
    arrange(desc(contig_freq_a))

  # Check the number of rows and consolidate into one row
  if (nrow(current_output) > 1) {
    # Create a new data frame with consolidated information
    consolidated_output <- data.frame(barcode = barcode)

    for (i in 1:nrow(current_output)) {
      # For each row, add its content to the consolidated output
      row_data <- current_output[i, -1]  # Exclude barcode column
      
      # Rename columns for all rows except the first one
      if (i > 1) {
        colnames(row_data) <- paste0(colnames(row_data), "_", i)
      }

      consolidated_output <- cbind(consolidated_output, row_data)
    }
    
    # Add the consolidated output to the final output data frame
    data_TRA_output <- bind_rows(data_TRA_output, consolidated_output)
  } else {
    # For single row barcodes, just add the current output
    data_TRA_output <- bind_rows(data_TRA_output, current_output)
  }
}
```


#### processing of data_TRB

```{r}
# Initialize an empty data frame to store the output
data_TRB_output <- data.frame()

# Get unique barcodes in the order they appear
unique_barcodes <- unique(data_TRB$barcode)

# Loop through each unique barcode
for (barcode in unique_barcodes) {
  # Filter the data for the current barcode
  current_data <- data_TRB %>%
    filter(barcode == !!barcode)

  # Count the number of rows for the current barcode, excluding the header
  total_contigs_b <- nrow(current_data)

  # Sum the reads column for the current barcode
  sum_reads <- sum(current_data$reads_b)

  # Calculate frequency of the current barcode
  contig_freq_b <- current_data$reads_b / sum_reads

  # Create a new data frame for the current barcode and move total_contigs column
  current_output <- current_data %>%
    mutate(total_contigs_b = total_contigs_b,
           contig_freq_b = contig_freq_b) %>%
    select(barcode, total_contigs_b, contig_freq_b, everything()) %>%
    arrange(desc(contig_freq_b))

  # Check the number of rows and consolidate into one row
  if (nrow(current_output) > 1) {
    # Create a new data frame with consolidated information
    consolidated_output <- data.frame(barcode = barcode)

    for (i in 1:nrow(current_output)) {
      # For each row, add its content to the consolidated output
      row_data <- current_output[i, -1]  # Exclude barcode column
      
      # Rename columns for all rows except the first one
      if (i > 1) {
        colnames(row_data) <- paste0(colnames(row_data), "_", i)
      }

      consolidated_output <- cbind(consolidated_output, row_data)
    }
    
    # Add the consolidated output to the final output data frame
    data_TRB_output <- bind_rows(data_TRB_output, consolidated_output)
  } else {
    # For single row barcodes, just add the current output
    data_TRB_output <- bind_rows(data_TRB_output, current_output)
  }
}
```

### Merge data_TRA_output and data_TRB_output with the barcodes object
```{r}
# Join barcodes with data_TRA_output
merged_TRA <- left_join(barcodes, data_TRA_output, by = "barcode")

# Join barcodes with data_TRB_output
merged_TRB <- left_join(barcodes, data_TRB_output, by = "barcode")

# Merge the two data frames based on the barcode column
merged_combined <- left_join(merged_TRA, merged_TRB, by = "barcode")
```

### reorder the columns to keep top contigs of TRA and TRB in close proximity. Finally export the merged_combined as an excel file
```{r}
# Get the column names for extensions "_a" and "_b"
cols_a <- grep("_a$", names(merged_combined), value = TRUE)
cols_b <- grep("_b$", names(merged_combined), value = TRUE)

# Get the remaining columns that do not have "_a" or "_b"
remaining_cols <- setdiff(names(merged_combined), c(cols_a, cols_b))

# Reorder the columns: first "_a", then "_b", followed by the remaining columns
# Include the 'barcode' column in the final selection
final_order <- c("barcode", cols_a, cols_b, remaining_cols)

# Reorganize the merged_combined data frame based on the final column order
merged_combined <- merged_combined %>%
  select(all_of(final_order))

# Export the merged_combined data frame to an Excel file in one line
write_xlsx(merged_combined, here("results", "merged_barcode_TRA_TRB.xlsx"))
```



------End------