---
title: "R Notebook - Custom gdVDJ Processing - Cellranger to Seurat compatible format"
output: html_notebook
---

#### Load required packages:
To load the required packages, execute the following code:

```{r}
library(tidyverse)
library(Matrix)
library(RCurl)
library(openxlsx)
library(here)
library(dplyr)
library(writexl)
```

Note: If you haven't installed the packages yet, please install them before loading.

#### Data loading
Load barcode.tsv.gz file from "cellranger count" output and "filtered_contig_annotations.csv" file (for ab T cells) or "raw_contig_annotations.csv" file (for gd T cells)
```{r}
# Read the .tsv.gz file that contains barcodes from the cellranger count output
barcodes <- read.delim(gzfile(here("data", "barcodes.tsv.gz")), header = FALSE, stringsAsFactors = FALSE)
colnames(barcodes) <- c("barcode")

# Read the CSV file
data <- read.csv(here("data", "all_contig_annotations.csv"), header = TRUE, stringsAsFactors = FALSE)
```

#### Split the filtered data into TRG and TRD (TRD and TRA)
```{r}

data_TRG <- data %>%
  filter(chain == "TRG")

data_TRD <- data %>%
  filter(chain == "TRD" | chain == "TRA")

# Rename columns in data_TRG while keeping 'barcode' unchanged
data_TRG <- data_TRG %>%
  rename_with(~ paste0(., "_g"), -barcode)  # Exclude 'barcode' from renaming

# Rename columns in data_TRD_or_TRA while keeping 'barcode' unchanged
data_TRD <- data_TRD %>%
  rename_with(~ paste0(., "_d"), -barcode)  # Exclude 'barcode' from renaming
```

#### processing of data_TRG and data_TRD

```{r}
######## processing of data_TRG #######

# Initialize an empty data frame to store the output
data_TRG_output <- data.frame()

# Get unique barcodes in the order they appear
unique_barcodes <- unique(data_TRG$barcode)

# Loop through each unique barcode
for (barcode in unique_barcodes) {
  # Filter the data for the current barcode
  current_data <- data_TRG %>%
    filter(barcode == !!barcode)

  # Count the number of rows for the current barcode, excluding the header
  total_contigs_g <- nrow(current_data)

  # Sum the reads column for the current barcode
  sum_reads <- sum(current_data$reads_g)

  # Calculate frequency of the current barcode
  contig_freq_g <- current_data$reads_g / sum_reads

  # Create a new data frame for the current barcode and move total_contigs column
  current_output <- current_data %>%
    mutate(total_contigs_g = total_contigs_g,
           contig_freq_g = contig_freq_g) %>%
    select(barcode, total_contigs_g, contig_freq_g, everything()) %>%
    arrange(desc(contig_freq_g))

  # Check the number of rows and consolidate into one row
  if (nrow(current_output) > 1) {
    # Create a new data frame with consolidated information
    consolidated_output <- data.frame(barcode = barcode)

    for (i in 1:nrow(current_output)) {
      # For each row, add its content to the consolidated output
      row_data <- current_output[i, -1]  # Exclude barcode column
      
      # Rename columns for all rows except the first one
      if (i > 1) {
        colnames(row_data) <- paste0(colnames(row_data), "_", i)
      }

      consolidated_output <- cbind(consolidated_output, row_data)
    }
    
    # Add the consolidated output to the final output data frame
    data_TRG_output <- bind_rows(data_TRG_output, consolidated_output)
  } else {
    # For single row barcodes, just add the current output
    data_TRG_output <- bind_rows(data_TRG_output, current_output)
  }
}

######## processing of data_TRD #######
# Initialize an empty data frame to store the output
data_TRD_output <- data.frame()

# Get unique barcodes in the order they appear
unique_barcodes <- unique(data_TRD$barcode)

# Loop through each unique barcode
for (barcode in unique_barcodes) {
  # Filter the data for the current barcode
  current_data <- data_TRD %>%
    filter(barcode == !!barcode)

  # Count the number of rows for the current barcode, excluding the header
  total_contigs_d <- nrow(current_data)

  # Sum the reads column for the current barcode
  sum_reads <- sum(current_data$reads_d)

  # Calculate frequency of the current barcode
  contig_freq_d <- current_data$reads_d / sum_reads

  # Create a new data frame for the current barcode and move total_contigs column
  current_output <- current_data %>%
    mutate(total_contigs_d = total_contigs_d,
           contig_freq_d = contig_freq_d) %>%
    select(barcode, total_contigs_d, contig_freq_d, everything()) %>%
    arrange(desc(contig_freq_d))

  # Check the number of rows and consolidate into one row
  if (nrow(current_output) > 1) {
    consolidated_output <- data.frame(barcode = barcode)

    for (i in 1:nrow(current_output)) {
      row_data <- current_output[i, -1]
      if (i > 1) {
        colnames(row_data) <- paste0(colnames(row_data), "_", i)
      }
      consolidated_output <- cbind(consolidated_output, row_data)
    }
    data_TRD_output <- bind_rows(data_TRD_output, consolidated_output)
  } else {
    data_TRD_output <- bind_rows(data_TRD_output, current_output)
  }
}

### Merge data_TRG_output and data_TRD_output with the barcodes object
# Join barcodes with data_TRG_output
merged_TRG <- left_join(barcodes, data_TRG_output, by = "barcode")

# Join barcodes with data_TRD_output
merged_TRD <- left_join(barcodes, data_TRD_output, by = "barcode")

# Merge the two data frames based on the barcode column
merged_combined <- left_join(merged_TRG, merged_TRD, by = "barcode")

### reorder the columns to keep top contigs of TRG and TRD/TRA in close proximity. Finally export the merged_combined as an excel file
# Get the column names for extensions "_g" and "_d"
cols_g <- grep("_g$", names(merged_combined), value = TRUE)
cols_d <- grep("_d$", names(merged_combined), value = TRUE)

# Get the remaining columns that do not have "_g" or "_d"
remaining_cols <- setdiff(names(merged_combined), c(cols_g, cols_d))

# Reorder the columns: first "_g", then "_d", followed by the remaining columns
final_order <- c("barcode", cols_g, cols_d, remaining_cols)

# Reorganize the merged_combined data frame based on the final column order
merged_combined <- merged_combined %>%
  select(all_of(final_order))

# Export the merged_combined data frame to an Excel file
write_xlsx(merged_combined, here("results", "merged_barcode_TRG_TRD.xlsx"))
```


------End------